---
title: "phenology_estimates_LM_JH"
output: html_document
date: '2023-08-03'
---
This script calculates the phenology estimates for B. philenor, L. a. astyanax, and L. archippus for the years 2019 - 2022 and latitudes 29º to 43º N. Rather than binning by latitude*year, this script only bins by latitude. We did this in order to increase data density for each quantile estimate and to compute more precise confidence intervals. Note that the phenology estimates calculated here will likely be different from the estimates for data binned by latitude AND year (the estimates used in our linear model).


Load packages
```{r}
library(curl)
library(ggplot2)
library(tidyverse)
library(dplyr)
library(lubridate)
library(sf)
library(spData)
library(phenesse)
```

Note - the data in my GitHub was downloaded from iNaturalist under the following parameters:
- quality grade: research
- reviewed: any
- identifications: most agree
- captive/cultivated: no
- date range: 01-01-2002 through 12-31-2022 (later pared down to 01-01-2015 through 12-31-2021)

```{r Battus data}
b <- curl("https://raw.githubusercontent.com/jnherr/iNaturalist-phenology-project/main/b_philenor_iNat_data.csv")
battus <- read.csv(b, header = TRUE, sep = ",")
head(battus)

length(battus$id) ### Sample size: 37984
b <- subset(battus, select = c(id, observed_on, latitude, longitude, scientific_name) )
head(b)

b <- b %>% filter(scientific_name %in% c("Battus philenor", "Battus philenor philenor")) #getting rid of other subspecies, e.g. B. p. hirsuta
length(b$id) # n = 36592
battus <- na.omit(b)
length(battus$id) ### New sample size: 36479

# Adding dates to whole Battus dataset
b.dates <- parse_date_time(battus$observed_on, "ymd")
b.months <- months(b.dates)
b.year <- year(b.dates)
b.day <- day(b.dates)
days_of_the_year <- yday(b.dates) #  days since Jan 1
battus$julian_day <- days_of_the_year
battus$year <- b.year
battus$month <- b.months
battus$day <- b.day
battus

# Adding longitudinal cutoff (see rationale for this cutoff in the plots made later in this document)
battus <- battus %>% filter(longitude > -102.5)
summary(battus)
battus
```

```{r L. a. astyanax data}
l <- curl("https://raw.githubusercontent.com/jnherr/iNaturalist-phenology-project/main/l_a_astyanax_iNat_data.csv")
astyanax <- read.csv(l, header = TRUE, sep = ",")
head(astyanax)

length(astyanax$id) ### Sample Size: 25148
l <- subset(astyanax, select = c(id, observed_on, latitude, longitude, scientific_name) )

l <- l %>% filter(scientific_name %in% "Limenitis arthemis astyanax") # getting rid of other subspecies (namely L. a. arizonensis)
length(l$id) # n = 24435
astyanax <- na.omit(l)
length(astyanax$id) ### New Sample Size: 24376

# Adding dates to whole Limenitis dataset
l.dates <- parse_date_time(astyanax$observed_on, "ymd")
l.months <- months(l.dates)
l.year <- year(l.dates)
l.day <- day(l.dates)
days_of_the_year <- yday(l.dates) #  days since Jan 1
astyanax$julian_day <- days_of_the_year
astyanax$year <- l.year
astyanax$month <- l.months
astyanax$day <- l.day
astyanax

# Adding longitudinal cutoff
astyanax <- astyanax %>% filter(longitude > -102.5)
summary(astyanax)
astyanax
```

```{r L. archippus data}
a <- curl("https://raw.githubusercontent.com/jnherr/iNaturalist-phenology-project/main/l_archippus_iNat_data.csv")
archippus <- read.csv(a, header = TRUE, sep = ",")
head(archippus)

length(archippus$id)
a <- subset(archippus, select = c(id, observed_on, latitude, longitude, scientific_name) )
head(a)

archippus <- na.omit(a)
length(archippus$id)

# Adding dates to whole archippus dataset
a.dates <- parse_date_time(archippus$observed_on, "ymd")
a.months <- months(a.dates)
a.year <- year(a.dates)
a.day <- day(a.dates)
days_of_the_year <- yday(a.dates) #  days since Jan 1
archippus$julian_day <- days_of_the_year
archippus$year <- a.year
archippus$month <- a.months
archippus$day <- a.day
archippus

# Adding longitudinal cutoff
archippus <- archippus %>% filter(longitude > -102.5)
summary(archippus)
archippus
```

```{r B. philenor by latitude}
# First subset all data by the total latitudinal range of interest.
battus <- battus %>%
  filter( latitude > "25", latitude < "43") %>%
  filter(year > 2018)
battus

# Next use the round() function to create a new column binning all latitudes into integer values.
battus$latitudinal_bin <- round(battus$latitude)
battus <- battus[order(battus$latitudinal_bin, decreasing=F),]
battus

# Check to ensure the data was binned correctly.
head(battus)
tail(battus) # looks good
length(battus$id) # n = 26364

# This next step checks the sample size for each latitudinal bin.

# 25ºN lat. bin
length(which(battus$latitudinal_bin==25))

# 26ºN lat. bin
length(which(battus$latitudinal_bin==26))

# 27ºN lat. bin
length(which(battus$latitudinal_bin==27))

# 28ºN lat. bin
length(which(battus$latitudinal_bin==28))

# 29ºN lat. bin
length(which(battus$latitudinal_bin==29))

# 30ºN lat. bin
length(which(battus$latitudinal_bin==30))

# 31ºN lat. bin
length(which(battus$latitudinal_bin==31))

# 32ºN lat. bin
length(which(battus$latitudinal_bin==32))

# 33ºN lat. bin
length(which(battus$latitudinal_bin==33))

# 34ºN lat. bin
length(which(battus$latitudinal_bin==34))

# 35ºN lat. bin
length(which(battus$latitudinal_bin==35))

# 36ºN lat. bin
length(which(battus$latitudinal_bin==36))

# 37ºN lat. bin
length(which(battus$latitudinal_bin==37))

# 38ºN lat. bin
length(which(battus$latitudinal_bin==38))

# 39ºN lat. bin
length(which(battus$latitudinal_bin==39))

# 40ºN lat. bin
length(which(battus$latitudinal_bin==40))

# 41ºN lat. bin
length(which(battus$latitudinal_bin==41))

# 42ºN lat. bin
length(which(battus$latitudinal_bin==42))

# 43ºN lat. bin
length(which(battus$latitudinal_bin==43)) # all binned sample counts exceed the minimum of 30 data points; looks good

# 44ºN lat. bin
length(which(battus$latitudinal_bin==44)) # insufficient data points

# 45ºN lat. bin
length(which(battus$latitudinal_bin==45)) # insufficient data points
```

```{r L. a. astyanax by latitude}
# First subset all data by the total latitudinal range of interest.
astyanax <- astyanax %>%
  filter( latitude > "25", latitude < "43") %>%
  filter(year > 2018)
astyanax

# Next use the round() function to create a new column binning all latitudes into integer values.
astyanax$latitudinal_bin <- round(astyanax$latitude)
astyanax <- astyanax[order(astyanax$latitudinal_bin, decreasing=F),]
astyanax

# Check to ensure the data was binned correctly.
head(astyanax)
tail(astyanax) # looks good
length(astyanax$id) # n = 21307

# This next step checks the sample size for each latitudinal bin.

# 25ºN lat. bin
length(which(astyanax$latitudinal_bin==25)) # insufficient data points

# 26ºN lat. bin
length(which(astyanax$latitudinal_bin==26)) # insufficient data points

# 27ºN lat. bin
length(which(astyanax$latitudinal_bin==27)) # insufficient data points

# 28ºN lat. bin
length(which(astyanax$latitudinal_bin==28)) # insufficient data points

# 29ºN lat. bin
length(which(astyanax$latitudinal_bin==29))  # this is where we start to get                                                    # sufficient data for phenesse
                                              # n = 65

# 30ºN lat. bin
length(which(astyanax$latitudinal_bin==30))

# 31ºN lat. bin
length(which(astyanax$latitudinal_bin==31))

# 32ºN lat. bin
length(which(astyanax$latitudinal_bin==32))

# 33ºN lat. bin
length(which(astyanax$latitudinal_bin==33))

# 34ºN lat. bin
length(which(astyanax$latitudinal_bin==34))

# 35ºN lat. bin
length(which(astyanax$latitudinal_bin==35))

# 36ºN lat. bin
length(which(astyanax$latitudinal_bin==36))

# 37ºN lat. bin
length(which(astyanax$latitudinal_bin==37))

# 38ºN lat. bin
length(which(astyanax$latitudinal_bin==38))

# 39ºN lat. bin
length(which(astyanax$latitudinal_bin==39))

# 40ºN lat. bin
length(which(astyanax$latitudinal_bin==40))

# 41ºN lat. bin
length(which(astyanax$latitudinal_bin==41))

# 42ºN lat. bin
length(which(astyanax$latitudinal_bin==42))

# 43ºN lat. bin
length(which(astyanax$latitudinal_bin==43)) # all binned sample counts exceed the minimum of 30 data points; also looks good
```

```{r Data filtering}
# First subset all data by the total latitudinal range of interest.
archippus <- archippus %>%
  filter( latitude > "25", latitude < "45") %>%
  filter(year > 2018)
archippus

# Next use the round() function to create a new column binning all latitudes into integer values.
archippus$latitudinal_bin <- round(archippus$latitude)
archippus <- archippus[order(archippus$latitudinal_bin, decreasing=F),]
archippus

# Check to ensure the data was binned correctly.
head(archippus)
tail(archippus) # looks good
length(archippus$id)

# This next step checks the sample size for each latitudinal bin.

# 29ºN lat. bin
length(which(archippus$latitudinal_bin==29))

# 30ºN lat. bin
length(which(archippus$latitudinal_bin==30))

# 31ºN lat. bin
length(which(archippus$latitudinal_bin==31))

# 32ºN lat. bin
length(which(archippus$latitudinal_bin==32))

# 33ºN lat. bin
length(which(archippus$latitudinal_bin==33))

# 34ºN lat. bin
length(which(archippus$latitudinal_bin==34))

# 35ºN lat. bin
length(which(archippus$latitudinal_bin==35))

# 36ºN lat. bin
length(which(archippus$latitudinal_bin==36))

# 37ºN lat. bin
length(which(archippus$latitudinal_bin==37))

# 38ºN lat. bin
length(which(archippus$latitudinal_bin==38))

# 39ºN lat. bin
length(which(archippus$latitudinal_bin==39))

# 40ºN lat. bin
length(which(archippus$latitudinal_bin==40))

# 41ºN lat. bin
length(which(archippus$latitudinal_bin==41))

# 42ºN lat. bin
length(which(archippus$latitudinal_bin==42))

# 43ºN lat. bin
length(which(archippus$latitudinal_bin==43))
```

```{r Capping binned obs. 29ºN}
# Latitudinal bin 29ºN
battus29 <- subset(battus, latitudinal_bin == "29")
astyanax29 <- subset(astyanax, latitudinal_bin == "29")
archippus29 <- subset(archippus, latitudinal_bin == "29")

# Now capping the number of observations per date (but not by Julian day/doy):
battus_capped29 = battus29[!duplicated(battus29$observed_on),]
battus_capped29[order(battus_capped29[,'julian_day']),]

astyanax_capped29 = astyanax29[!duplicated(astyanax29$observed_on),]
astyanax_capped29[order(astyanax_capped29[,'julian_day']),]

archippus_capped29 = archippus29[!duplicated(archippus29$observed_on),]
archippus_capped29[order(archippus_capped29[,'julian_day']),]

battus_capped29 # n = 720
astyanax_capped29 # n = 62
archippus_capped29 # n = 339
```

```{r Capping binned obs. 30ºN}
battus30 <- subset(battus, latitudinal_bin == "30")
astyanax30 <- subset(astyanax, latitudinal_bin == "30")
archippus30 <- subset(archippus, latitudinal_bin == "30")

battus_capped30 = battus30[!duplicated(battus30$observed_on),]
battus_capped30[order(battus_capped30[,'julian_day']),]

astyanax_capped30 = astyanax30[!duplicated(astyanax30$observed_on),]
astyanax_capped30[order(astyanax_capped30[,'julian_day']),]

archippus_capped30 = archippus30[!duplicated(archippus30$observed_on),]
archippus_capped30[order(archippus_capped30[,'julian_day']),]

battus_capped30 # n = 1069
astyanax_capped30 # n = 311
archippus_capped30 # n = 518
```

```{r Capping binned obs. 31ºN}
battus31 <- subset(battus, latitudinal_bin == "31")
astyanax31 <- subset(astyanax, latitudinal_bin == "31")
archippus31 <- subset(archippus, latitudinal_bin == "31")

battus_capped31 = battus31[!duplicated(battus31$observed_on),]
battus_capped31[order(battus_capped31[,'julian_day']),]

astyanax_capped31 = astyanax31[!duplicated(astyanax31$observed_on),]
astyanax_capped31[order(astyanax_capped31[,'julian_day']),]

archippus_capped31 = archippus31[!duplicated(archippus31$observed_on),]
archippus_capped31[order(archippus_capped31[,'julian_day']),]

battus_capped31 # n = 564
astyanax_capped31 # n = 261
archippus_capped31 # n = 266
```

```{r Capping binned obs. 32ºN}
battus32 <- subset(battus, latitudinal_bin == "32")
astyanax32 <- subset(astyanax, latitudinal_bin == "32")
archippus32 <- subset(archippus, latitudinal_bin == "32")

battus_capped32 = battus32[!duplicated(battus32$observed_on),]
battus_capped32[order(battus_capped32[,'julian_day']),]

astyanax_capped32 = astyanax32[!duplicated(astyanax32$observed_on),]
astyanax_capped32[order(astyanax_capped32[,'julian_day']),]

archippus_capped32 = archippus32[!duplicated(archippus32$observed_on),]
archippus_capped32[order(archippus_capped32[,'julian_day']),]

battus_capped32 # n = 301
astyanax_capped32 # n = 204
archippus_capped32 # n = 196
```

```{r Capping binned obs. 33ºN}
battus33 <- subset(battus, latitudinal_bin == "33")
astyanax33 <- subset(astyanax, latitudinal_bin == "33")
archippus33 <- subset(archippus, latitudinal_bin == "33")

battus_capped33 = battus33[!duplicated(battus33$observed_on),]
battus_capped33[order(battus_capped33[,'julian_day']),]

astyanax_capped33 = astyanax33[!duplicated(astyanax33$observed_on),]
astyanax_capped33[order(astyanax_capped33[,'julian_day']),]

archippus_capped33 = archippus33[!duplicated(archippus33$observed_on),]
archippus_capped33[order(archippus_capped33[,'julian_day']),]

battus_capped33 # n = 689
astyanax_capped33 # n = 417
archippus_capped33 # n = 422
```

```{r Capping binned obs. 34ºN}
battus34 <- subset(battus, latitudinal_bin == "34")
astyanax34 <- subset(astyanax, latitudinal_bin == "34")
archippus34 <- subset(archippus, latitudinal_bin == "34")

battus_capped34 = battus34[!duplicated(battus34$observed_on),]
battus_capped34[order(battus_capped34[,'julian_day']),]

astyanax_capped34 = astyanax34[!duplicated(astyanax34$observed_on),]
astyanax_capped34[order(astyanax_capped34[,'julian_day']),]

archippus_capped34 = archippus34[!duplicated(archippus34$observed_on),]
archippus_capped34[order(archippus_capped34[,'julian_day']),]

battus_capped34 # n = 317
astyanax_capped34 # n = 516
archippus_capped34 # n = 229
```

```{r Capping binned obs. 35ºN}
battus35 <- subset(battus, latitudinal_bin == "35")
astyanax35 <- subset(astyanax, latitudinal_bin == "35")
archippus35 <- subset(archippus, latitudinal_bin == "35")

battus_capped35 = battus35[!duplicated(battus35$observed_on),]
battus_capped35[order(battus_capped35[,'julian_day']),]

astyanax_capped35 = astyanax35[!duplicated(astyanax35$observed_on),]
astyanax_capped35[order(astyanax_capped35[,'julian_day']),]

archippus_capped35 = archippus35[!duplicated(archippus35$observed_on),]
archippus_capped35[order(archippus_capped35[,'julian_day']),]

battus_capped35 # n = 616
astyanax_capped35 # n = 574
archippus_capped35 # n = 251
```

```{r Capping binned obs. 36ºN}
battus36 <- subset(battus, latitudinal_bin == "36")
astyanax36 <- subset(astyanax, latitudinal_bin == "36")
archippus36 <- subset(archippus, latitudinal_bin == "36")

battus_capped36 = battus36[!duplicated(battus36$observed_on),]
battus_capped36[order(battus_capped36[,'julian_day']),]

astyanax_capped36 = astyanax36[!duplicated(astyanax36$observed_on),]
astyanax_capped36[order(astyanax_capped36[,'julian_day']),]

archippus_capped36 = archippus36[!duplicated(archippus36$observed_on),]
archippus_capped36[order(archippus_capped36[,'julian_day']),]

battus_capped36 # n = 726
astyanax_capped36 # n = 637
archippus_capped36 # n = 375
```


```{r Capping binned obs. 37ºN}
battus37 <- subset(battus, latitudinal_bin == "37")
astyanax37 <- subset(astyanax, latitudinal_bin == "37")
archippus37 <- subset(archippus, latitudinal_bin == "37")

battus_capped37 = battus37[!duplicated(battus37$observed_on),]
battus_capped37[order(battus_capped37[,'julian_day']),]

astyanax_capped37 = astyanax37[!duplicated(astyanax37$observed_on),]
astyanax_capped37[order(astyanax_capped37[,'julian_day']),]

archippus_capped37 = archippus37[!duplicated(archippus37$observed_on),]
archippus_capped37[order(archippus_capped37[,'julian_day']),]

battus_capped37 # n = 386
astyanax_capped37 # n = 419
archippus_capped37 # n = 206
```

```{r Capping binned obs. 38ºN}
battus38 <- subset(battus, latitudinal_bin == "38")
astyanax38 <- subset(astyanax, latitudinal_bin == "38")
archippus38 <- subset(archippus, latitudinal_bin == "38")

battus_capped38 = battus38[!duplicated(battus38$observed_on),]
battus_capped38[order(battus_capped38[,'julian_day']),]

astyanax_capped38 = astyanax38[!duplicated(astyanax38$observed_on),]
astyanax_capped38[order(astyanax_capped38[,'julian_day']),]

archippus_capped38 = archippus38[!duplicated(archippus38$observed_on),]
archippus_capped38[order(archippus_capped38[,'julian_day']),]

battus_capped38 # n = 424
astyanax_capped38 # n = 464
archippus_capped38 # n = 203
```

```{r Capping binned obs. 39ºN}
battus39 <- subset(battus, latitudinal_bin == "39")
astyanax39 <- subset(astyanax, latitudinal_bin == "39")
archippus39 <- subset(archippus, latitudinal_bin == "39")

battus_capped39 = battus39[!duplicated(battus39$observed_on),]
battus_capped39[order(battus_capped39[,'julian_day']),]

astyanax_capped39 = astyanax39[!duplicated(astyanax39$observed_on),]
astyanax_capped39[order(astyanax_capped39[,'julian_day']),]

archippus_capped39 = archippus39[!duplicated(archippus39$observed_on),]
archippus_capped39[order(archippus_capped39[,'julian_day']),]

battus_capped39 # n = 507
astyanax_capped39 # n = 591
archippus_capped39 # n = 447
```

```{r Capping binned obs. 40ºN}
battus40 <- subset(battus, latitudinal_bin == "40")
astyanax40 <- subset(astyanax, latitudinal_bin == "40")
archippus40 <- subset(archippus, latitudinal_bin == "40")

battus_capped40 = battus40[!duplicated(battus40$observed_on),]
battus_capped40[order(battus_capped40[,'julian_day']),]

astyanax_capped40 = astyanax40[!duplicated(astyanax40$observed_on),]
astyanax_capped40[order(astyanax_capped40[,'julian_day']),]

archippus_capped40 = archippus40[!duplicated(archippus40$observed_on),]
archippus_capped40[order(archippus_capped40[,'julian_day']),]

battus_capped40 # n = 235
astyanax_capped40 # n = 523
archippus_capped40 # n = 367
```

```{r Capping binned obs. 41ºN}
battus41 <- subset(battus, latitudinal_bin == "41")
astyanax41 <- subset(astyanax, latitudinal_bin == "41")
archippus41 <- subset(archippus, latitudinal_bin == "41")

battus_capped41 = battus41[!duplicated(battus41$observed_on),]
battus_capped41[order(battus_capped41[,'julian_day']),]

astyanax_capped41 = astyanax41[!duplicated(astyanax41$observed_on),]
astyanax_capped41[order(astyanax_capped41[,'julian_day']),]

archippus_capped41 = archippus41[!duplicated(archippus41$observed_on),]
archippus_capped41[order(archippus_capped41[,'julian_day']),]

battus_capped41 # n = 167
astyanax_capped41 # n = 430
archippus_capped41 # n = 388
```

```{r Capping binned obs. 42ºN}
battus42 <- subset(battus, latitudinal_bin == "42")
astyanax42 <- subset(astyanax, latitudinal_bin == "42")
archippus42 <- subset(archippus, latitudinal_bin == "42")

battus_capped42 = battus42[!duplicated(battus42$observed_on),]
battus_capped42[order(battus_capped42[,'julian_day']),]

astyanax_capped42 = astyanax42[!duplicated(astyanax42$observed_on),]
astyanax_capped42[order(astyanax_capped42[,'julian_day']),]

archippus_capped42 = archippus42[!duplicated(archippus42$observed_on),]
archippus_capped42[order(archippus_capped42[,'julian_day']),]

battus_capped42 # n = 82
astyanax_capped42 # n = 389
archippus_capped42 # n = 457
```

```{r Capping binned obs. 43ºN}
battus43 <- subset(battus, latitudinal_bin == "43")
astyanax43 <- subset(astyanax, latitudinal_bin == "43")
archippus43 <- subset(archippus, latitudinal_bin == "43")

battus_capped43 = battus43[!duplicated(battus43$observed_on),]
battus_capped43[order(battus_capped43[,'julian_day']),]

astyanax_capped43 = astyanax43[!duplicated(astyanax43$observed_on),]
astyanax_capped43[order(astyanax_capped43[,'julian_day']),]

archippus_capped43 = archippus43[!duplicated(archippus43$observed_on),]
archippus_capped43[order(archippus_capped43[,'julian_day']),]

battus_capped43 # n = 34
astyanax_capped43 # n = 262
archippus_capped43 # n = 450
```

```{r 29ºN quantile estimates}
quantile_ci(observations = battus_capped29$julian_day, percentile = 0.10, bootstraps = 10000, conf = 0.95, type = "perc")
quantile_ci(observations = astyanax_capped29$julian_day, percentile = 0.10, bootstraps = 10000, conf = 0.95, type = "perc")
quantile_ci(observations = archippus_capped29$julian_day, percentile = 0.10, bootstraps = 10000, conf = 0.95, type = "perc")

quantile_ci(observations = battus_capped29$julian_day, percentile = 0.90, bootstraps = 10000, conf = 0.95, type = "perc")
quantile_ci(observations = astyanax_capped29$julian_day, percentile = 0.90, bootstraps = 10000, conf = 0.95, type = "perc")
quantile_ci(observations = archippus_capped29$julian_day, percentile = 0.90, bootstraps = 10000, conf = 0.95, type = "perc")
```

```{r 30ºN quantile estimates}
quantile_ci(observations = battus_capped30$julian_day, percentile = 0.10, bootstraps = 10000, conf = 0.95, type = "perc")
quantile_ci(observations = astyanax_capped30$julian_day, percentile = 0.10, bootstraps = 10000, conf = 0.95, type = "perc")
quantile_ci(observations = archippus_capped30$julian_day, percentile = 0.10, bootstraps = 10000, conf = 0.95, type = "perc")

quantile_ci(observations = battus_capped30$julian_day, percentile = 0.90, bootstraps = 10000, conf = 0.95, type = "perc")
quantile_ci(observations = astyanax_capped30$julian_day, percentile = 0.90, bootstraps = 10000, conf = 0.95, type = "perc")
quantile_ci(observations = archippus_capped30$julian_day, percentile = 0.90, bootstraps = 10000, conf = 0.95, type = "perc")
```

```{r 31ºN quantile estimates}
quantile_ci(observations = battus_capped31$julian_day, percentile = 0.10, bootstraps = 10000, conf = 0.95, type = "perc")
quantile_ci(observations = astyanax_capped31$julian_day, percentile = 0.10, bootstraps = 10000, conf = 0.95, type = "perc")
quantile_ci(observations = archippus_capped31$julian_day, percentile = 0.10, bootstraps = 10000, conf = 0.95, type = "perc")

quantile_ci(observations = battus_capped31$julian_day, percentile = 0.90, bootstraps = 10000, conf = 0.95, type = "perc")
quantile_ci(observations = astyanax_capped31$julian_day, percentile = 0.90, bootstraps = 10000, conf = 0.95, type = "perc")
quantile_ci(observations = archippus_capped31$julian_day, percentile = 0.90, bootstraps = 10000, conf = 0.95, type = "perc")
```

```{r 32ºN quantile estimates}
quantile_ci(observations = battus_capped32$julian_day, percentile = 0.10, bootstraps = 10000, conf = 0.95, type = "perc")
quantile_ci(observations = astyanax_capped32$julian_day, percentile = 0.10, bootstraps = 10000, conf = 0.95, type = "perc")
quantile_ci(observations = archippus_capped32$julian_day, percentile = 0.10, bootstraps = 10000, conf = 0.95, type = "perc")

quantile_ci(observations = battus_capped32$julian_day, percentile = 0.90, bootstraps = 10000, conf = 0.95, type = "perc")
quantile_ci(observations = astyanax_capped32$julian_day, percentile = 0.90, bootstraps = 10000, conf = 0.95, type = "perc")
quantile_ci(observations = archippus_capped32$julian_day, percentile = 0.90, bootstraps = 10000, conf = 0.95, type = "perc")
```

```{r 33ºN quantile estimates}
quantile_ci(observations = battus_capped33$julian_day, percentile = 0.10, bootstraps = 10000, conf = 0.95, type = "perc")
quantile_ci(observations = astyanax_capped33$julian_day, percentile = 0.10, bootstraps = 10000, conf = 0.95, type = "perc")
quantile_ci(observations = archippus_capped33$julian_day, percentile = 0.10, bootstraps = 10000, conf = 0.95, type = "perc")

quantile_ci(observations = battus_capped33$julian_day, percentile = 0.90, bootstraps = 10000, conf = 0.95, type = "perc")
quantile_ci(observations = astyanax_capped33$julian_day, percentile = 0.90, bootstraps = 10000, conf = 0.95, type = "perc")
quantile_ci(observations = archippus_capped33$julian_day, percentile = 0.90, bootstraps = 10000, conf = 0.95, type = "perc")
```

```{r 34ºN quantile estimates}
quantile_ci(observations = battus_capped34$julian_day, percentile = 0.10, bootstraps = 10000, conf = 0.95, type = "perc")
quantile_ci(observations = astyanax_capped34$julian_day, percentile = 0.10, bootstraps = 10000, conf = 0.95, type = "perc")
quantile_ci(observations = archippus_capped34$julian_day, percentile = 0.10, bootstraps = 10000, conf = 0.95, type = "perc")

quantile_ci(observations = battus_capped34$julian_day, percentile = 0.90, bootstraps = 10000, conf = 0.95, type = "perc")
quantile_ci(observations = astyanax_capped34$julian_day, percentile = 0.90, bootstraps = 10000, conf = 0.95, type = "perc")
quantile_ci(observations = archippus_capped34$julian_day, percentile = 0.90, bootstraps = 10000, conf = 0.95, type = "perc")
```

```{r 35ºN quantile estimates}
quantile_ci(observations = battus_capped35$julian_day, percentile = 0.10, bootstraps = 10000, conf = 0.95, type = "perc")
quantile_ci(observations = astyanax_capped35$julian_day, percentile = 0.10, bootstraps = 10000, conf = 0.95, type = "perc")
quantile_ci(observations = archippus_capped35$julian_day, percentile = 0.10, bootstraps = 10000, conf = 0.95, type = "perc")

quantile_ci(observations = battus_capped35$julian_day, percentile = 0.90, bootstraps = 10000, conf = 0.95, type = "perc")
quantile_ci(observations = astyanax_capped35$julian_day, percentile = 0.90, bootstraps = 10000, conf = 0.95, type = "perc")
quantile_ci(observations = archippus_capped35$julian_day, percentile = 0.90, bootstraps = 10000, conf = 0.95, type = "perc")
```

```{r 36ºN quantile estimates}
quantile_ci(observations = battus_capped36$julian_day, percentile = 0.10, bootstraps = 10000, conf = 0.95, type = "perc")
quantile_ci(observations = astyanax_capped36$julian_day, percentile = 0.10, bootstraps = 10000, conf = 0.95, type = "perc")
quantile_ci(observations = archippus_capped36$julian_day, percentile = 0.10, bootstraps = 10000, conf = 0.95, type = "perc")

quantile_ci(observations = battus_capped36$julian_day, percentile = 0.90, bootstraps = 10000, conf = 0.95, type = "perc")
quantile_ci(observations = astyanax_capped36$julian_day, percentile = 0.90, bootstraps = 10000, conf = 0.95, type = "perc")
quantile_ci(observations = archippus_capped36$julian_day, percentile = 0.90, bootstraps = 10000, conf = 0.95, type = "perc")
```

```{r 37ºN quantile estimates}
quantile_ci(observations = battus_capped37$julian_day, percentile = 0.10, bootstraps = 10000, conf = 0.95, type = "perc")
quantile_ci(observations = astyanax_capped37$julian_day, percentile = 0.10, bootstraps = 10000, conf = 0.95, type = "perc")
quantile_ci(observations = archippus_capped37$julian_day, percentile = 0.10, bootstraps = 10000, conf = 0.95, type = "perc")

quantile_ci(observations = battus_capped37$julian_day, percentile = 0.90, bootstraps = 10000, conf = 0.95, type = "perc")
quantile_ci(observations = astyanax_capped37$julian_day, percentile = 0.90, bootstraps = 10000, conf = 0.95, type = "perc")
quantile_ci(observations = archippus_capped37$julian_day, percentile = 0.90, bootstraps = 10000, conf = 0.95, type = "perc")
```

```{r 38ºN quantile estimates}
quantile_ci(observations = battus_capped38$julian_day, percentile = 0.10, bootstraps = 10000, conf = 0.95, type = "perc")
quantile_ci(observations = astyanax_capped38$julian_day, percentile = 0.10, bootstraps = 10000, conf = 0.95, type = "perc")
quantile_ci(observations = archippus_capped38$julian_day, percentile = 0.10, bootstraps = 10000, conf = 0.95, type = "perc")

quantile_ci(observations = battus_capped38$julian_day, percentile = 0.90, bootstraps = 10000, conf = 0.95, type = "perc")
quantile_ci(observations = astyanax_capped38$julian_day, percentile = 0.90, bootstraps = 10000, conf = 0.95, type = "perc")
quantile_ci(observations = archippus_capped38$julian_day, percentile = 0.90, bootstraps = 10000, conf = 0.95, type = "perc")
```

```{r 39ºN quantile estimates}
quantile_ci(observations = battus_capped39$julian_day, percentile = 0.10, bootstraps = 10000, conf = 0.95, type = "perc")
quantile_ci(observations = astyanax_capped39$julian_day, percentile = 0.10, bootstraps = 10000, conf = 0.95, type = "perc")
quantile_ci(observations = archippus_capped39$julian_day, percentile = 0.10, bootstraps = 10000, conf = 0.95, type = "perc")

quantile_ci(observations = battus_capped39$julian_day, percentile = 0.90, bootstraps = 10000, conf = 0.95, type = "perc")
quantile_ci(observations = astyanax_capped39$julian_day, percentile = 0.90, bootstraps = 10000, conf = 0.95, type = "perc")
quantile_ci(observations = archippus_capped39$julian_day, percentile = 0.90, bootstraps = 10000, conf = 0.95, type = "perc")
```

```{r 40ºN quantile estimates}
quantile_ci(observations = battus_capped40$julian_day, percentile = 0.10, bootstraps = 10000, conf = 0.95, type = "perc")
quantile_ci(observations = astyanax_capped40$julian_day, percentile = 0.10, bootstraps = 10000, conf = 0.95, type = "perc")
quantile_ci(observations = archippus_capped40$julian_day, percentile = 0.10, bootstraps = 10000, conf = 0.95, type = "perc")

quantile_ci(observations = battus_capped40$julian_day, percentile = 0.90, bootstraps = 10000, conf = 0.95, type = "perc")
quantile_ci(observations = astyanax_capped40$julian_day, percentile = 0.90, bootstraps = 10000, conf = 0.95, type = "perc")
quantile_ci(observations = archippus_capped40$julian_day, percentile = 0.90, bootstraps = 10000, conf = 0.95, type = "perc")
```

```{r 41ºN quantile estimates}
quantile_ci(observations = battus_capped41$julian_day, percentile = 0.10, bootstraps = 10000, conf = 0.95, type = "perc")
quantile_ci(observations = astyanax_capped41$julian_day, percentile = 0.10, bootstraps = 10000, conf = 0.95, type = "perc")
quantile_ci(observations = archippus_capped41$julian_day, percentile = 0.10, bootstraps = 10000, conf = 0.95, type = "perc")

quantile_ci(observations = battus_capped41$julian_day, percentile = 0.90, bootstraps = 10000, conf = 0.95, type = "perc")
quantile_ci(observations = astyanax_capped41$julian_day, percentile = 0.90, bootstraps = 10000, conf = 0.95, type = "perc")
quantile_ci(observations = archippus_capped41$julian_day, percentile = 0.90, bootstraps = 10000, conf = 0.95, type = "perc")
```

```{r 42ºN quantile estimates}
quantile_ci(observations = battus_capped42$julian_day, percentile = 0.10, bootstraps = 10000, conf = 0.95, type = "perc")
quantile_ci(observations = astyanax_capped42$julian_day, percentile = 0.10, bootstraps = 10000, conf = 0.95, type = "perc")
quantile_ci(observations = archippus_capped42$julian_day, percentile = 0.10, bootstraps = 10000, conf = 0.95, type = "perc")

quantile_ci(observations = battus_capped42$julian_day, percentile = 0.90, bootstraps = 10000, conf = 0.95, type = "perc")
quantile_ci(observations = astyanax_capped42$julian_day, percentile = 0.90, bootstraps = 10000, conf = 0.95, type = "perc")
quantile_ci(observations = archippus_capped42$julian_day, percentile = 0.90, bootstraps = 10000, conf = 0.95, type = "perc")
```

```{r 43ºN quantile estimates}
quantile_ci(observations = battus_capped43$julian_day, percentile = 0.10, bootstraps = 10000, conf = 0.95, type = "perc")
quantile_ci(observations = astyanax_capped43$julian_day, percentile = 0.10, bootstraps = 10000, conf = 0.95, type = "perc")
quantile_ci(observations = archippus_capped43$julian_day, percentile = 0.10, bootstraps = 10000, conf = 0.95, type = "perc")

quantile_ci(observations = battus_capped43$julian_day, percentile = 0.90, bootstraps = 10000, conf = 0.95, type = "perc")
quantile_ci(observations = astyanax_capped43$julian_day, percentile = 0.90, bootstraps = 10000, conf = 0.95, type = "perc")
quantile_ci(observations = archippus_capped43$julian_day, percentile = 0.90, bootstraps = 10000, conf = 0.95, type = "perc")
```


