---
title: "pheno_plots_JH"
output: html_document
date: '2023-08-03'
---
Script for creating phenology plots including all three species: B. philenor, L. a. astyanax, and L. archippus.

```{r Load packages}
library(curl)
library(ggplot2)
library(tidyverse)
library(dplyr)
library(spData)
library(sf)
library(Rmisc)
```

```{r Template map of the US}
x_extent <- c(-90, -70)
y_extent <- c(25, 43)

sf_northamerica <- world |> 
  filter(continent == "North America") |>
  group_by(continent) |>
  tally()

sf_northamerica <- ggplot(data = sf_northamerica) +
  geom_sf() +
  coord_sf(xlim = x_extent, ylim = y_extent) +
  theme_classic()

sf_northamerica
```

Loading the 2019 - 2022 filtered dataframe (created in Excel, using Excel functions to calculate onset and offset lag times and overlap of yearly flight times per latitudinal bin). This dataset was binned only by latitude, and the quantile estimates were calculated for each latitudinal bin (four years of data per bin).

```{r 2019 - 2022 butterfly data}
QE_lm1_df <- curl("https://raw.githubusercontent.com/jnherr/iNaturalist-phenology-project/main/phenology_LM1_cumulative_df.csv")
QE_lm1_df <- read.csv(QE_lm1_df, header = TRUE, sep = ",")
QE_lm1_df
````

```{r LM cumulative maps}
# With error bars
onset_all1 <- ggplot(data = QE_lm1_df, aes(x = onset_doy, y = lat, color = species)) +
  geom_point(size = 2) +
  geom_errorbar(aes(x = onset_doy, y = lat, xmax = onset_high_ci, xmin = onset_low_ci)) +
  scale_color_manual(name = "Species",
                     breaks = c('b_philenor', 'la_astyanax', 'l_archippus'),
                     values = c("blue","skyblue3","orange")) +
  labs(title = "Avg. onset estimates", color = "Species") +
  xlab("Day of Year") +
  ylab("Latitude (ºN)") +
  theme_classic()

# With line segments
onset_all2 <- ggplot(data = QE_lm1_df, aes(x = onset_doy, y = lat, color = species)) +
  geom_point(size = 2) +
  geom_segment(aes(x = onset_low_ci, y = lat, xend = onset_high_ci, yend = lat)) +
  scale_color_manual(name = "Species",
                     breaks = c('b_philenor', 'la_astyanax', 'l_archippus'),
                     values = c("blue","skyblue3","orange")) +
  labs(title = "Avg. onset estimates", color = "Species") +
  xlab("Day of Year") +
  ylab("Latitude (ºN)") +
  theme_classic()
  
onset_all1
onset_all2
```

```{r LM Battus-astyanax map}
# B. philenor and L. a. astyanax
bat_ast_df <- subset(QE_lm1_df, !(species %in% c("l_archippus")))

onset_bat_ast <- ggplot(data = bat_ast_df, aes(x = onset_doy, y = lat, color = species)) +
  geom_point(size = 2) +
  xlim(0, 200) +
  # geom_errorbar(aes(x = onset_doy, y = lat, xmax = onset_high_ci, xmin = onset_low_ci)) +
  geom_segment(aes(x = onset_low_ci, y = lat, xend = onset_high_ci, yend = lat)) +
  scale_color_manual(name = "Species",
                     breaks = c('b_philenor', 'la_astyanax'),
                     values = c("blue","skyblue3")) +
  # labs(title = "Avg. onset estimates - B. philenor and L. a. astyanax", color = "Species") +
  xlab("Day of Year") +
  ylab("Latitude (ºN)") +
  theme_classic()

onset_bat_ast
```

```{r LM Battus-archippus plot}
# B. philenor and L. archippus
bat_arch_df <- subset(QE_lm1_df, !(species %in% c("la_astyanax")))

onset_bat_arch <- ggplot(data = bat_arch_df, aes(x = onset_doy, y = lat, color = species)) +
  geom_point(size = 2) +
  xlim(0, 200) +
  # geom_errorbar(aes(x = onset_doy, y = lat, xmax = onset_high_ci, xmin = onset_low_ci)) +
  geom_segment(aes(x = onset_low_ci, y = lat, xend = onset_high_ci, yend = lat)) +
  scale_color_manual(name = "Species",
                     breaks = c('b_philenor', 'l_archippus'),
                     values = c("blue","orange")) +
  # labs(title = "Avg. onset estimates - B. philenor and L. archippus", color = "Species") +
  xlab("Day of Year") +
  ylab("Latitude (ºN)") +
  theme_classic()

onset_bat_arch
```

Next step: plotting data for each year, 2019 - 2022 (data binned by latitude*year). Previously, I had averaged the quantile estimates across all four years, but now I'm interested in seeing if year has any noticeable effect on the phenesse onset estimates.

```{r LMER butterfly data}
QE_lmer_df <- curl("https://raw.githubusercontent.com/jnherr/iNaturalist-phenology-project/main/phenology_LMER_df.csv")
QE_lmer_df <- read.csv(QE_lmer_df, header = TRUE, sep = ",")
QE_lmer_df
```

```{r 2019 onset estimate plots}
onset_2019_df <- subset(QE_lmer_df, (year %in% c("2019")))

# With error bars
onset_2019 <- ggplot(data = onset_2019_df, aes(x = onset_doy, y = lat, color = species)) +
  geom_point(size = 2) +
  xlim(0, 250) +
  geom_errorbar(aes(x = onset_doy, y = lat, xmax = onset_high_ci, xmin = onset_low_ci)) +
  scale_color_manual(name = "Species",
                     breaks = c('b_philenor', 'la_astyanax', 'l_archippus'),
                     values = c("blue","skyblue3","orange")) +
  labs(title = "Avg. onset estimates (2019)", color = "Species") +
  xlab("Day of Year") +
  ylab("Latitude (ºN)") +
  theme_classic()

# B. philenor and L. a. astyanax
bat_ast_2019 <- subset(onset_2019_df, !(species %in% c("l_archippus")))

onset_bat_ast_2019 <- ggplot(data = bat_ast_2019, aes(x = onset_doy, y = lat, color = species)) +
  geom_point(size = 2) +
  xlim(0, 250) +
  #geom_errorbar(aes(x = onset_doy, y = lat, xmax = onset_high_ci, xmin = onset_low_ci)) +
  geom_segment(aes(x = onset_low_ci, y = lat, xend = onset_high_ci, yend = lat)) +
  scale_color_manual(name = "Species",
                     breaks = c('b_philenor', 'la_astyanax'),
                     values = c("blue","skyblue3")) +
  labs(title = "Avg. onset estimates - B. philenor and L. a. astyanax (2019)", color = "Species") +
  xlab("Day of Year") +
  ylab("Latitude (ºN)") +
  theme_classic()

# B. philenor and L. archippus
bat_arch_2019 <- subset(onset_2019_df, !(species %in% c("la_astyanax")))

onset_bat_arch_2019 <- ggplot(data = bat_arch_2019, aes(x = onset_doy, y = lat, color = species)) +
  geom_point(size = 2) +
  xlim(0, 250) +
  #geom_errorbar(aes(x = onset_doy, y = lat, xmax = onset_high_ci, xmin = onset_low_ci)) +
  geom_segment(aes(x = onset_low_ci, y = lat, xend = onset_high_ci, yend = lat)) +
  scale_color_manual(name = "Species",
                     breaks = c('b_philenor', 'l_archippus'),
                     values = c("blue","orange")) +
  labs(title = "Avg. onset estimates - B. philenor and L. archippus (2019)", color = "Species") +
  xlab("Day of Year") +
  ylab("Latitude (ºN)") +
  theme_classic()



onset_2019
onset_bat_ast_2019
onset_bat_arch_2019
```

```{r 2020 onset estimate plots}
onset_2020_df <- subset(QE_lmer_df, (year %in% c("2020")))

# With error bars
onset_2020 <- ggplot(data = onset_2020_df, aes(x = onset_doy, y = lat, color = species)) +
  geom_point(size = 2) +
  xlim(0, 250) +
  geom_errorbar(aes(x = onset_doy, y = lat, xmax = onset_high_ci, xmin = onset_low_ci)) +
  scale_color_manual(name = "Species",
                     breaks = c('b_philenor', 'la_astyanax', 'l_archippus'),
                     values = c("blue","skyblue3","orange")) +
  labs(title = "Avg. onset estimates (2020)", color = "Species") +
  xlab("Day of Year") +
  ylab("Latitude (ºN)") +
  theme_classic()

# B. philenor and L. a. astyanax
bat_ast_2020 <- subset(onset_2020_df, !(species %in% c("l_archippus")))

onset_bat_ast_2020 <- ggplot(data = bat_ast_2020, aes(x = onset_doy, y = lat, color = species)) +
  geom_point(size = 2) +
  xlim(0, 250) +
  #geom_errorbar(aes(x = onset_doy, y = lat, xmax = onset_high_ci, xmin = onset_low_ci)) +
  geom_segment(aes(x = onset_low_ci, y = lat, xend = onset_high_ci, yend = lat)) +
  scale_color_manual(name = "Species",
                     breaks = c('b_philenor', 'la_astyanax'),
                     values = c("blue","skyblue3")) +
  labs(title = "Avg. onset estimates - B. philenor and L. a. astyanax (2020)", color = "Species") +
  xlab("Day of Year") +
  ylab("Latitude (ºN)") +
  theme_classic()

# B. philenor and L. archippus
bat_arch_2020 <- subset(onset_2020_df, !(species %in% c("la_astyanax")))

onset_bat_arch_2020 <- ggplot(data = bat_arch_2020, aes(x = onset_doy, y = lat, color = species)) +
  geom_point(size = 2) +
  xlim(0, 250) +
  #geom_errorbar(aes(x = onset_doy, y = lat, xmax = onset_high_ci, xmin = onset_low_ci)) +
  geom_segment(aes(x = onset_low_ci, y = lat, xend = onset_high_ci, yend = lat)) +
  scale_color_manual(name = "Species",
                     breaks = c('b_philenor', 'l_archippus'),
                     values = c("blue","orange")) +
  labs(title = "Avg. onset estimates - B. philenor and L. archippus (2020)", color = "Species") +
  xlab("Day of Year") +
  ylab("Latitude (ºN)") +
  theme_classic()



onset_2020
onset_bat_ast_2020
onset_bat_arch_2020
```

```{r 2021 onset estimate plots}
onset_2021_df <- subset(QE_lmer_df, (year %in% c("2021")))

# With error bars
onset_2021 <- ggplot(data = onset_2021_df, aes(x = onset_doy, y = lat, color = species)) +
  geom_point(size = 2) +
  xlim(0, 200) +
  geom_errorbar(aes(x = onset_doy, y = lat, xmax = onset_high_ci, xmin = onset_low_ci)) +
  scale_color_manual(name = "Species",
                     breaks = c('b_philenor', 'la_astyanax', 'l_archippus'),
                     values = c("blue","skyblue3","orange")) +
  labs(title = "Avg. onset estimates (2021)", color = "Species") +
  xlab("Day of Year") +
  ylab("Latitude (ºN)") +
  theme_classic()

# B. philenor and L. a. astyanax
bat_ast_2021 <- subset(onset_2021_df, !(species %in% c("l_archippus")))

onset_bat_ast_2021 <- ggplot(data = bat_ast_2021, aes(x = onset_doy, y = lat, color = species)) +
  geom_point(size = 2) +
  xlim(0, 200) +
  #geom_errorbar(aes(x = onset_doy, y = lat, xmax = onset_high_ci, xmin = onset_low_ci)) +
  geom_segment(aes(x = onset_low_ci, y = lat, xend = onset_high_ci, yend = lat)) +
  scale_color_manual(name = "Species",
                     breaks = c('b_philenor', 'la_astyanax'),
                     values = c("blue","skyblue3")) +
  labs(title = "Avg. onset estimates - B. philenor and L. a. astyanax (2021)", color = "Species") +
  xlab("Day of Year") +
  ylab("Latitude (ºN)") +
  theme_classic()

# B. philenor and L. archippus
bat_arch_2021 <- subset(onset_2021_df, !(species %in% c("la_astyanax")))

onset_bat_arch_2021 <- ggplot(data = bat_arch_2021, aes(x = onset_doy, y = lat, color = species)) +
  geom_point(size = 2) +
  xlim(0, 200) +
  #geom_errorbar(aes(x = onset_doy, y = lat, xmax = onset_high_ci, xmin = onset_low_ci)) +
  geom_segment(aes(x = onset_low_ci, y = lat, xend = onset_high_ci, yend = lat)) +
  scale_color_manual(name = "Species",
                     breaks = c('b_philenor', 'l_archippus'),
                     values = c("blue","orange")) +
  labs(title = "Avg. onset estimates - B. philenor and L. archippus (2021)", color = "Species") +
  xlab("Day of Year") +
  ylab("Latitude (ºN)") +
  theme_classic()



onset_2021
onset_bat_ast_2021
onset_bat_arch_2021
```

```{r 2022 onset estimate plots}
onset_2022_df <- subset(QE_lmer_df, (year %in% c("2022")))

# With error bars
onset_2022 <- ggplot(data = onset_2022_df, aes(x = onset_doy, y = lat, color = species)) +
  geom_point(size = 2) +
  xlim(0, 200) +
  geom_errorbar(aes(x = onset_doy, y = lat, xmax = onset_high_ci, xmin = onset_low_ci)) +
  scale_color_manual(name = "Species",
                     breaks = c('b_philenor', 'la_astyanax', 'l_archippus'),
                     values = c("blue","skyblue3","orange")) +
  labs(title = "Avg. onset estimates (2022)", color = "Species") +
  xlab("Day of Year") +
  ylab("Latitude (ºN)") +
  theme_classic()

# B. philenor and L. a. astyanax
bat_ast_2022 <- subset(onset_2022_df, !(species %in% c("l_archippus")))

onset_bat_ast_2022 <- ggplot(data = bat_ast_2022, aes(x = onset_doy, y = lat, color = species)) +
  geom_point(size = 2) +
  xlim(0, 200) +
  #geom_errorbar(aes(x = onset_doy, y = lat, xmax = onset_high_ci, xmin = onset_low_ci)) +
  geom_segment(aes(x = onset_low_ci, y = lat, xend = onset_high_ci, yend = lat)) +
  scale_color_manual(name = "Species",
                     breaks = c('b_philenor', 'la_astyanax'),
                     values = c("blue","skyblue3")) +
  labs(title = "Avg. onset estimates - B. philenor and L. a. astyanax (2022)", color = "Species") +
  xlab("Day of Year") +
  ylab("Latitude (ºN)") +
  theme_classic()

# B. philenor and L. archippus
bat_arch_2022 <- subset(onset_2022_df, !(species %in% c("la_astyanax")))

onset_bat_arch_2022 <- ggplot(data = bat_arch_2022, aes(x = onset_doy, y = lat, color = species)) +
  geom_point(size = 2) +
  xlim(0, 200) +
  #geom_errorbar(aes(x = onset_doy, y = lat, xmax = onset_high_ci, xmin = onset_low_ci)) +
  geom_segment(aes(x = onset_low_ci, y = lat, xend = onset_high_ci, yend = lat)) +
  scale_color_manual(name = "Species",
                     breaks = c('b_philenor', 'l_archippus'),
                     values = c("blue","orange")) +
  labs(title = "Avg. onset estimates - B. philenor and L. archippus (2022)", color = "Species") +
  xlab("Day of Year") +
  ylab("Latitude (ºN)") +
  theme_classic()



onset_2022
onset_bat_ast_2022
onset_bat_arch_2022
```

```{r Avg. annual onset dates per species}
avg_onset_df <- QE_lmer_df %>%
  select(onset_doy, year, species) %>%
  group_by(species, year) %>%
  mutate(species_annual_onset = mean(onset_doy)) %>%
  ungroup() %>%
  select(species_annual_onset, year, species) %>%
  group_by(year) %>%
  distinct()

avg_onset_df    # "species_annual_onset" is the annual avg. onset time per species per year
```

```{r Plot comparing annual onsets by spp.}
avg_onset_plot <- ggplot(data = avg_onset_df, aes(x = year, y = species_annual_onset, color = species)) +
  geom_line(size = 1) +
  scale_color_manual(name = "Species",
                     breaks = c("b_philenor", "la_astyanax", "l_archippus"),
                     values = c("blue","lightblue", "orange")) +
  # labs(title = "Avg. annual onset estimates", color = "Species") +
  xlab("Year") +
  ylab("Estimated onset day") +
  theme_classic()

avg_onset_plot
```

```{r Calculating avg. annual onset w/ SD, SE, CI}
avg_onsets = summarySE(data = QE_lmer_df, measurevar = "onset_doy", groupvars = c("year","species"))

avg_onsets # Here, "onset_doy" is the mean onset date by spp*yr.
```

```{r Box and whisker plots of avg. onset}
avg_onset_plot2 <- ggplot(data = avg_onsets, aes(x = year, y = onset_doy)) +
  geom_point(aes(color = species), size = 4) +
  geom_errorbar(aes(x = year, y = onset_doy, ymax = onset_doy + ci, ymin = onset_doy - ci, color = species), width = 0.2) +
  scale_color_manual(name = "Species",
                    breaks = c("b_philenor", "la_astyanax", "l_archippus"),
                    values = c("blue","lightblue", "orange")) +
  xlab("Year") +
  ylab("Estimated onset day") +
  theme_classic()

avg_onset_plot2
```
